---
title: "week2"
author: "Ilir_Sheraj"
date: "8/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Vectors

**Vector**: The most basic unit for data storage

```{r}
codes <- c(380, 124, 818)
country <- c("italy", "canada", "egypt")
```

Name entries of vector

```{r}
codes <- c(italy = 380, canada = 124, egypt = 818)
codes
```

```{r}
class(codes)
```

We can do the same by using quotes to define the names.

```{r}
codes <- c(380, 124, 818)
country <- c("italy", "canada", "egypt")
names(codes) <- country
codes
```

# Sequences

```{r}
seq(1,10)
```

```{r}
seq(1,10, 2)
```

```{r}
1:10
```

# Subsetting

```{r}
codes[2]
```

```{r}
codes[c(1,3)]
```

```{r}
codes[1:2]
```

```{r}
codes["canada"]
```

```{r}
codes[c("canada", "egypt")]
```


# Vector Coercion

**Coercion** An attempt by R to be flexible with data types

```{r}
x <- c(1, "canada", 3)
x
```

```{r}
x <- 1:5
y <- as.character(x)
y
```

```{r}
z <- as.numeric(y)
z
```

Coercion can introduce NAs

```{r}
x <- c("1", "b", "3")
as.numeric(x)
```

# Assessment

Use the function c() to create a numeric vector with the average high temperatures in January for Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto. The average high temperatures are 35, 88, 42, 84, 81, and 30 degrees Fahrenheit. Call the object temp.

```{r}
temp <- c("Beijing" = 35, "Lagos" = 88, "Paris" = 42, "Rio de Janeiro" = 84,
                  "San Juan" = 81,"Toronto" = 30)
temp
```

The temperatures we stored in temp are from the cities Beijing, Lagos, Paris, Rio de Janeiro, San Juan, and Toronto. Create a vector with these city names and call the object city. Make sure to use the correct order, spelling and capitalization!

```{r}
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan","Toronto")
city
```

Use the names function and the objects defined in the previous exercises to associate the temperature data with its corresponding city. (You can go back to the previous questions and copy the objects stored.) Note: to see what happened, after assigning the city names to the temp vector, try printing the temp vector to understand how the names are associated with elements of temp.

```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
names(temp) <- city
temp
```

Use the [ and : operators to access the temperature of the first three cities in the list, which are already stored in temp.

```{r}
temp[1:3]
```

This time we will access our object using just the [ operator. Use the [ operator to access the temperature of Paris and San Juan already stored in the temp object.

```{r}
temp[c(3,5)]
```

This time we will use just the : operator. Use the : operator to create a sequence of consecutive integers starting at 12 and ending at 73 and save it in an object x, then determine the length of object x.

```{r}
x <- 12:73
length(x)
```

We will learn how to use the seq() function in this question. Create a vector containing all the positive odd numbers smaller than 100. The numbers should be in ascending order.

```{r}
seq(1, 100, 2)
```

Create a vector of numbers that starts at 6, does not go beyond 55, and adds numbers in increments of 4/7. So the first three numbers will be 6, 6+4/7, and 6+8/7. How many numbers does the list have? Use only one line of code to answer both questions.

```{r}
length(seq(6, 55, 4/7))
```

Determine the class of a vector generated with seq using the length.out argument. Specifically, what is the class of the following object a <- seq(1, 10, length.out = 100)?

```{r}
a <- seq(1, 10, length.out = 100)
class(a)
```

The main difference between numeric and integer is that integers occupy less space in the computer memory, so for big computations using integers can have a substantial impact. Does this change depending on what we store in an object? What is the class of the following object a <- seq(1, 10)?

```{r}
a <- seq(1, 10)
class(a)
a
```

Define the following vector: x <- c(1, 3, 5,"a"). Notice that x has been coerced into a character string. If we assign something new to a vector that we've already created, the new definition replaces whatever was previously assigned to the vector. Convert x to a vector of numbers using the as.numeric() function. (Note that you will get a warning message, but that is okay!)

```{r}
x <- c(1, 3, 5, "a")
x <- as.numeric(x)
x
```

# Sorting

We will deal with gunmurders data

```{r}
library(dslabs)
data("murders")
sort(murders$total)
```

```{r}
x <- c(31, 4, 15, 92, 65)
sort(x)
```

Order gives us the index

```{r}
index <- order(x)
index
x[index]
```

```{r}
index <- order(murders$total)
murders$abb[index]
```

```{r}
max(murders$total)
```

```{r}
i_max <- which.max(murders$total)
murders$state[i_max]
```

```{r}
i_min <- which.min(murders$total)
murders$state[i_min]
```

```{r}
x <- c(31, 4, 15, 92, 65)
rank(x)
```

# Assessment

Use the $ operator to access the population size data and store it in the object pop. Then use the sort function to redefine pop so that it is sorted. Finally use the [ operator to report the smallest population size.

```{r}
pop <- murders$population
pop <- sort(pop)
pop[1]
```

Now instead of the smallest population size, let's find out the row number, in the data frame murders, of the state with the smallest population size. This time we need to use order() instead of sort(). Remember that the entries in the vector murders$population follow the order of the rows of murders.

```{r}
pop <- murders$population
ord <- order(murders$population)
ord
```

Write one line of code that gives the index of the lowest population entry. Use the which.min command.

```{r}
which.min(murders$population)
```

Find the index of the smallest state using which.min(murders$population). Define a variable states to hold the state names from the murders data frame. Combine these to find the state name for the smallest state.

```{r}
i <- which.min(murders$population)
states <- murders$state
states[i]
```

You can create a data frame using the data.frame function. Here is a quick example:

```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
city_temps <- data.frame(name = city, temperature = temp)
city_temps
```

Define a variable states to be the state names from murders Use rank(murders$population) to determine the population size rank (from smallest to biggest) of each state. Save these ranks in an object called ranks.Create a data frame with state names and their respective ranks. Call the data frame my_df.

```{r}
states <- murders$state
ranks <- rank(murders$population)
my_df <- data.frame(states, ranks)
head(my_df)
```

Create variables states and ranks to store the state names and ranks by population size respectively. Create an object ind that stores the indexes needed to order the population values, using the order command. For example we could define o <- order(murders$population) Create a data frame with both variables following the correct order. Use the bracket operator [ to re-order each column in the data frame. For example, states[o] orders the abbreviations based by population size. The columns of the data frame must be in the specific order and have the specific names: states, ranks.

```{r}
states <- murders$state
ranks <- rank(murders$population)
ind <- order(murders$population)
my_df <- data.frame(states[ind], ranks[ind])
head(my_df)
```

The na_example dataset represents a series of counts. It is included in the dslabs package. You can quickly examine the object using

```{r}
library(dslabs)
data("na_example")
str(na_example)
```

However, when we compute the average we obtain an NA. You can see this by typing

```{r}
mean(na_example)
```

The is.na returns a logical vector that tells us which entries are NA. Assign the logical vector that is returned by is.na(na_example) to an object called ind. Determine how many NAs na_example has, using the sum command.

```{r}
ind <- is.na(na_example)
sum(ind)
```

Write one line of code to compute the average, but only for the entries that are not NA making use of the ! operator before ind. (Remember that you can use help("!") to find out more about what ! does.)

```{r}
mean(na_example[!ind])
```

# Vectorized Arithmetic

```{r}
murders$state[which.max(murders$population)]
```

```{r}
max(murders$population)
```

Murder per capita

```{r}
heights <- c(69, 62, 66, 70, 70, 73, 67, 73, 67, 70)
heights*2.54
```

```{r}
heights - 69
```

```{r}
murder_rate <- murders$total/murders$population*100000
murder_rate
```

```{r}
murders$state[order(murder_rate, decreasing = TRUE)]
```

# Assessment

Previously we created this data frame:

```{r}
temp <- c(35, 88, 42, 84, 81, 30)
city <- c("Beijing", "Lagos", "Paris", "Rio de Janeiro", "San Juan", "Toronto")
city_temps <- data.frame(name = city, temperature = temp)
```

We are interested in specifying Celsius instead of Fahrenheit. Remember that to convert from Fahrenheit to Celsius we use c = 5/9 * (F - 32). Use vector arithmetic to convert temp to Celsius Create a data frame called city_temps with the city names and temperatures in Celsius.

```{r}
temp <- 5/9 * (temp - 32)
city_temps <- data.frame(name = city, temperature = temp)
city_temps
```

What is the sum of the following equation: 1 + 1/2^2 + 1/3^2 + ... + 1/100^2? Thanks to Euler we know it should be close to pi^2/6 Define an object x that contains the numbers 1 through 100 Compute the sum 1 + 1/2^2 + 1/3^2 + ... + 1/100^2.

```{r}
x <- 1:100
sum(1/(x^2))
```

Compute the per 100,000 murder rate for each state and store it in the object murder_rate. Then compute the average of the state murder rates for the US using the function mean. What is this average?

```{r}
murder_rate <- murders$total/murders$population*100000
mean(murder_rate)
```

# Week Assessment

```{r}
x <- c(2, 43, 27, 96, 18)
sort(x)
```

```{r}
order(x)
```

```{r}
rank(x)
```

```{r}
min(x)
```

```{r}
which.min(x)
```

```{r}
max(x)
```

```{r}
which.max(x)
```

Mandi, Amy, Nicole, and Olivia all ran different distances in different time intervals. Their distances (in miles) and times (in minutes) are as follows:

```{r}
name <- c("Mandi", "Amy", "Nicole", "Olivia")
distance <- c(0.8, 3.1, 2.8, 4.0)
time <- c(10, 30, 40, 50)
```

Write a line of code to convert time to hours. Remember there are 60 minutes in an hour. Then write a line of code to calculate the speed of each runner in miles per hour. Speed is distance divided by time. How many hours did Olivia run?

```{r}
time_h <- time/60
speed <- distance/time_h
runs <- data.frame(name = name, time = time_h, speed = speed)
runs
```

