---
title: "week4"
author: "Ilir_Sheraj"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Conditional Statements

```{r}
a <- 0
if (a != 0){
  print(1/a)
}else{
  print("There is no reciprocl of a")
}
```

```{r}
library(dslabs)
data("murders")
rate <- murders$total/murders$population*100000
ind <- which.min(rate)
if (rate[ind] < 0.5){
  print(murders$state[ind])
}else{
  print("No State has murder rat ethat low")
}
```

```{r}
a <- 0
ifelse(a !=0, 1/a, NA)
```

The power of ifelse is in vectors as it examines each element of the vector and returns an answer accordingly.

```{r}
a <- c(0, 1, 2, -4, 5)
result <- ifelse(a > 0, 1/a, NA)
result
```

Replacing NAs with any other value

```{r}
data("na_example")
sum(is.na(na_example))
```

```{r}
no_nas <- ifelse(is.na(na_example), 0, na_example)
sum(is.na(no_nas))
```

any() function takes a logical and returns true if any of the entries fulfills the condition

```{r}
z <- c(TRUE, TRUE, FALSE)
any(z)
```

```{r}
z <- c(FALSE, FALSE, FALSE)
any(z)
```

```{r}
z <- c(TRUE, TRUE, FALSE)
all(z)
```

```{r}
z <- c(TRUE, TRUE, TRUE)
all(z)
```

# Functions

```{r}
a <- 1:10
avg <- function(x){
  s <- sum(x)
  l <- length(x)
  s/l
}
avg(a)
```

```{r}
avg <- function(x, arithmetic = TRUE){
  n <- length(x)
  ifelse(arithmetic, sum(x)/n, prod(x)^(1/n))
}
```

# For Loops

```{r}
compute_s_n <- function(n){
  x <- 1:n
  sum(x)
}
compute_s_n(3)
```

```{r}
compute_s_n(1000)
```

```{r}
for (i in 1:5){
  print(i)
}
```

```{r}
m <- 25
s_n <- vector("numeric", length = m)
for (i in 1:m){
  s_n[i] <- compute_s_n(i)
}
s_n
```

```{r}
n <- 1:25
plot(n, s_n)
```

```{r}
plot(n, s_n)
lines(n, n*(n+1)/2)
```

# Assessment

What will this conditional expression return? Run it from the console.

```{r}
x <- c(1,2,-3,4)
if(all(x > 0)){
  print("All Positives")
}else{
  print("Not all Positives")
}
```

The function nchar tells you how many characters long a character vector is. For example:
```{r}
library(dslabs)
data("murders")
char_len <- nchar(murders$state)
head(char_len)
```

The function ifelse is useful because you convert a vector of logicals into something else. For example, some datasets use the number -999 to denote NA. A bad practice! You can convert the -999 in a vector to NA using the following ifelse call:

```{r}
x <- c(2, 3, -999, 1, 4, 5, -999, 3, 2, 9)
ifelse(x == -999, NA, x)
```

We will combine a number of functions for this exercise. Use the ifelse function to write one line of code that assigns to the object new_names the state abbreviation when the state name is longer than 8 characters and assigns the state name when the name is not longer than 8 characters. For example, where the original vector has Massachusetts (13 characters), the new vector should have MA. But where the original vector has New York (8 characters), the new vector should have New York as well.

```{r}
new_names <- ifelse(nchar(murders$state) > 8, murders$abb, murders$state)
```

You will encounter situations in which the function you need does not already exist. R permits you to write your own. Let's practice one such situation, in which you first need to define the function to be used. The functions you define can have multiple arguments as well as default values. To define functions we use function. For example the following function adds 1 to the number it receives as an argument:

```{r}
my_func <- function(x){
  y <- x + 1
  y
}
my_func(4)
```

We will define a function sum_n for this exercise. Create a function sum_n that for any given value, say n, creates the vector 1:n, and then computes the sum of the integers from 1 to n. Use the function you just defined to determine the sum of integers from 1 to 5,000.

```{r}
sum_n <- function(n){
  x <- 1:n
  sum(x)
}
sum_n(5000)
```

We will make another function for this exercise. We will define a function altman_plot that takes two arguments x and y and plots the difference y-x in the y-axis against the sum x+y in the x-axis. You can define functions with as many variables as you want. For example, here we need at least two, x and y. The following function plots log transformed values:

```{r}
log_plot <- function(x, y){
  plot(log10(x), log10(y))
}
log_plot(1:20, 1:20)
```

We will make another function for this exercise. Create a function altman_plot that takes two arguments x and y and plots y-x (on the y-axis) against x+y (on the x-axis). Note: don't use parentheses around the arguments in the plot function because you will confuse R.

```{r}
altman_plot <- function(x,y){
  plot(x+y, y-x)
}
altman_plot(1:10, 1:10)
```

Lexical scoping is a convention used by many languages that determine when an object is available by its name. When you run the code below you will see which x is available at different points in the code. Note that when we define x as 9, this is inside the function, but it is 8 after you run the function. The x changed inside the function but not outside.

```{r}
x <- 8
my_func <- function(y){
  x <- 9
  print(x)
  y + x
}
my_func(x)
print(x)
```

After running the code below, what is the value of x?

```{r}
x <- 3
my_func <- function(y){
  x <- 5
  y
  print(x)
}
my_func(x)
x
```

In the next exercise we are going to write a for-loop. In that for-loop we are going to call a function. We define that function here. Write a function compute_s_n that for any given n computes the sum 1^2 + 2^2 + ... + n^2

```{r}
n <- 10
compute_s_n <- function(n){
  x <- 1:n
  y <- x^2
  sum(y)
}
compute_s_n(10)
```

Now we are going to compute the sum of the squares for several values of n. . We will use a for-loop for this. Here is an example of a for-loop:

```{r}
results <- vector("numeric", 10)
n <- 10
for (i in 1:n){
  x <- 1:i
  results[i] <- sum(x)
}
results
```

Define an empty numeric vector s_n of size 25 using s_n <- vector("numeric", 25). Compute the the sum when n is equal to each integer from 1 to 25 using the function we defined in the previous exercise: compute_s_n


```{r}
s_n <- vector("numeric", 25)
n <- 1:25
for (i in n){
  s_n[i] <- compute_s_n(i)
}
s_n
```

Plot the results

```{r}
plot(n, s_n)
```

Conform whether the values are the same as expected mathematically

```{r}
identical(s_n, (n*(n+1)*(2*n+1))/6)
```

# Week Assessment

```{r}
data("heights")
head(heights)
```

Write an ifelse() statement that returns 1 if the sex is Female and 2 if the sex is Male. What is the sum of the returning vector?

```{r}
sum(ifelse(heights$sex=="Female", 1, 2))
```

Write an ifelse() statement that takes the height column and returns the height if it is greater than 72 inches and returns 0 otherwise. What is the mean of the resulting vector?

```{r}
mean(ifelse(heights$height > 72, heights$height, 0))
```

Write a function inches_to_ft that takes a number of inches x and returns the number of feet. One foot equals 12 inches. What is inches_to_ft(144)?

```{r}
inches_to_feet <- function(x){
  x*(1/12)
}
inches_to_feet(144)
```

How many individuals in the heights dataset have a height less than 5 feet?

```{r}
sum(heights$height < 60)
```

Factorial: Complete the code above to generate a vector of length m where the first entry is 1!, the second entry is 2!, and so on up to m!.

```{r}
m <- 10
f_n <- vector(length = m)
for (n in 1:m){
  f_n[n] <- factorial(n)
}
f_n
```

```{r}
plot(1:m, f_n)
lines(1:m, f_n)
```

